Network Configuration Synthesis with Abstract Topologies \vspace{1cm}

### 2017-03-15 paper/abstraction/paper.tex
more typo fixes and edits to paper

At a higher level, what is happening is that each inference rule is attempting to learn the maximum possible fault tolerance information possible as a function of the symbolic inputs, and $\nu$Z~\cite{z3opt} will minimize this maximum by accounting for all possible topologies that meet the abstraction. Facts learned with $j=0$ or $k=0$ are discarded.

@@

For example, in the concrete PG in Figure~\ref{fig:example-compilation}, router S1 appears in two PG nodes: $(S1,1,0)$ and $(S1,0,1)$. These nodes have two peers that can send them messages: $(N1,0,0)$ and $(N2,0,0)$. Therefore, S1 will allow advertisements from both peer $N1$ and peer $N2$. If S1 uses the path advertised by neighbor N1, then it will add the tag $(1,0)$ to the advertisement before sending this to its neighbors A1 and A2. If S1 uses a path advertised from neighbor N2, then it will add the tag $(0,1)$ instead. Similarly, router A1 appears in two separate PG nodes. It will allow advertisements from S1 if they either have the $(1,0)$ tag or the $(0,1)$ tag attached. In either case, it will not modify the tag before exporting the route to its neighbors.

@@

  %\vspace{1em}%
  \begin{minipage}[t]{.47\linewidth}
  \hdr{mBGP Syntax}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
     d    &\in& \textit{Integers} & \\
     c    &\in& \textit{Communities} & \\
     l    &\in& \textit{Topology Locations} & \\
     t    &::=& \$x \BNFALT d.d.d.d/[d..d] & \textit{predicate} \\
     ns   &::=& \{ l_1, ~\dots,~ l_k \} & \textit{peers} \\
     ma   &::=& d : ({ns}_1, c_1) \rightarrow ({ns}_2, c_2) & \textit{match action} \\
     pc   &::=& ma_1, \dots, ma_k & \textit{prefix config} \\
     rc   &::=& t_1 \rightarrow {pc}_1, ~\dots,~ t_k \rightarrow {pc}_k & \textit{router config} \\
     mbgp &::=& l_1 \rightarrow {rc}_1, ~\dots,~ l_k \rightarrow {rc}_k & \textit{mbgp policy} \\%
  \end{array} \]%%

@@

The idea then is to search for such a device-local preference function for each router that totally orders advertisements from different neighbors (possibly with different tags). For example, to satisfy the policy that leaving through N1 is preferred to leaving through N2, S1 should prefer a message it hears from N1 over N2. Similarly, A1 should prefer routes tagged with $(1,0)$ over those with $(0,1)$. Intuitively, S1 should prefer a message from N1 because it results in an accepting state for automata 1, which indicates a better path. Further, it can result in downstream routers such as A1 using a path that is accepting for automata 1 as well.

@@

\begin{defn}
We write $m \leq n$ if the subgraph reachable from $m$ and $n$ respectively form a simulation relation with respect to $\geq_{rank}$.
\end{defn}

More specifically, we say that $m \leq n$ if $m \geq_{rank} n$ and for every transition $n\overset{l}{\rightarrow}n'$ from PG node $n$ there exists a transition $m\overset{l}{\rightarrow}m'$ from $m$ and $m' \leq n'$.
%
If $m \leq n$, then advertisements received for node $m$ can safely be preferred over those received for node $n$ after accounting for the network-wide impact of the choice. For example in Figure~\ref{fig:example-compilation}, S1 can receive advertisements in two different contexts ($(S1,1,0)$ and $(S1,0,1)$) and must choose between messages received in these different contexts. Advertisements are preferred in state $(S1,1,0)$ because it results in a better path for S1 ($(S1,0,1) \geq_{rank} (S1,1,0)$) and downstream routers such as A1 will also obtain paths no worse than if S1 had choosen $(S1,1,0)$ instead. The policy is guaranteed to be safe from failures because if a link fails in the topology, then $m \leq n$ will still hold since any transition that becomes unusable for $m$ also becomes unusable for $n$. That is $m \leq n$ before the failure implies $m \leq n$ after the failure.

For each router, its corresponding PG nodes are then sorted according the the $\leq$ operator. If the operator defines a total order on PG nodes, then the compiler can simply prefer advertisements from peers of node $m$ over those of $n$ whenever $n \leq m$. However, if $\leq$ does not form a total order, then the policy is rejected as being potentially unsafe under some failure conditions.

@@

Figure~\ref{fig:concretization} (right) defines compilation from \sysname to \mbgp. It proceeds by compiling constraints $p_i$ in the original \sysname policy to a tuple of: the predicate $t_i$, the product graph $PG_i$, and the preference function $pref_i$. These tuples are passed the the $\mathrm{compile}_\mathrm{mBGP}$ function, along with the network topology $G$. For each internal router in the topology $l$, and each predicate $t_i$ in the \sysname policy, compilation goes through each node $m$ for $l$ in $PG_i$, and groups the inbound neighbors of $m$ by tag ($q_n$) into sets ($in$). It allows imports from these neighbors before exporting to the outbound neighbors of $m$. The local preference for these imports is given by $\mathrm{pref}_i(m)$. We build configs using list-comprehension notation. For instance, $[~ l \rightarrow rc ~\vert~ l \in V, p(l,rc) ~]$ denotes the \mbgp policy $l_1 \rightarrow rc_{1},...,l_k \rightarrow rc_{k}$ where each $rc_{i}$ satisfies $p(l_i,rc_{i})$. We also use $\mathit{append}$ to denote sequence concatenation.%

@@

\para{Concretization}
The similar structure between the spine template and concrete configurations is not a coincidence.  We formalize this observation by defining a two concretization function ($con$), one for \sysname policies and another for \mbgp policies. 
%
Concretization takes a context $\Gamma : \mathit{Var} \rightarrow 2^{\mathit{Prefix} \times V}$ that maps each template variable to a set of pairs of a concrete prefix and topology location where the prefix is owned.

Both concretization functions traverse the policy and substitute each instance of a topology loction $l$ in the template policy with the set of all concrete locations that map to $l$, given by the inverse homomorphism $f^{-1}(l) = \{l' ~\vert~ f(l') = l \}$. Additionally, whenver a pair $(pfx,l) \in \Gamma(x)$, then a new entry is added to the concretized policy where $pfx$ replaces $\$\mathrm{x}$ and adds the constraint that traffic ends at $l$ ($\CD{\End(l)}$).
%
For example, the spine template in Figure~\ref{fig:bgp-configs}, is obtained by substituting $\{ A1,A2 \}$ for $AG$ and $\{ A3,A4 \}$ for $AL$ and by also replacing the entry for {\small \$GP} with entries for {\small GP1} and {\small GP2} given by the context (\EG, a database).

>>>>

At a higher level, what is happening is that each inference rule is attempting to learn the maximum fault tolerance information possible as a function of the symbolic inputs, and $\nu$Z~\cite{z3opt} will minimize this maximum by accounting for all possible topologies that meet the abstraction. Facts learned with $j=0$ or $k=0$ are discarded.

@@

For example, in the concrete PG in Figure~\ref{fig:example-compilation}, router S1 appears in two PG nodes: $(S1,1,0)$ and $(S1,0,1)$. These nodes have two peers that can send them messages: $(N1,0,0)$ and $(N2,0,0)$. Therefore, S1 will allow advertisements from both peer $N1$ and peer $N2$. If S1 uses the path advertised by neighbor N1, then it will add the tag $(1,0)$ to the advertisement before sending this to its neighbors A1 and A2. If S1 uses a path advertised from neighbor N2, then it will add the tag $(0,1)$ instead. Similarly, router A1 appears in two PG nodes. It will allow advertisements from that have the $(1,0)$ or $(0,1)$ tag attached. In either case, it will not modify the tag before exporting the route to its neighbors.

@@

  %\vspace{1em}%
  \begin{minipage}[t]{.47\linewidth}
  \hdr{mBGP Syntax}{}
  \vspace*{-1\baselineskip}
  %
  \[ \begin{array}{rclr}
     d    &\in& \textit{Integers} & \\
     c    &\in& \textit{Communities} & \\
     l    &\in& \textit{Topology Locations} & \\
     t    &::=& \$x \BNFALT d.d.d.d/[d..d] & \textit{predicate} \\
     ns   &::=& \{ l_1, ~\dots,~ l_k \} & \textit{peers} \\
     ma   &::=& d : ({ns}_1, c_1) \rightarrow ({ns}_2, c_2) & \textit{match action} \\
     pc   &::=& ma_1, \dots, ma_k & \textit{predicate config} \\
     rc   &::=& t_1 \rightarrow {pc}_1, ~\dots,~ t_k \rightarrow {pc}_k & \textit{router config} \\
     mbgp &::=& l_1 \rightarrow {rc}_1, ~\dots,~ l_k \rightarrow {rc}_k & \textit{mbgp policy} \\%
  \end{array} \]%%

@@

The idea is to search for such a device-local preference function for each router that totally orders advertisements from different neighbors (possibly with different tags). For example, to satisfy the policy that leaving through N1 is preferred to leaving through N2, S1 should prefer a message it hears from N1 over N2. Similarly, A1 should prefer routes tagged with $(1,0)$ over those with $(0,1)$. Intuitively, S1 should prefer a message from N1 because it results in an accepting state for automata 1, which indicates a better path. Further, it can result in downstream routers such as A1 using a path that is accepting for automata 1 as well.

@@

\begin{defn}
We write $m \leq n$ if the subgraph reachable from $m$ and $n$ respectively form a simulation relation with respect to $\geq_{rank}$. More specifically, we say that $m \leq n$ if $n \geq_{rank} m$ and for every transition $n\overset{l}{\rightarrow}n'$ from PG node $n$ there exists a transition $m\overset{l}{\rightarrow}m'$ from $m$ and $m' \leq n'$.
\end{defn}

If $m \leq n$, then advertisements received for node $m$ can safely be preferred over those received for node $n$ after accounting for the network-wide impact of the choice. For example in Figure~\ref{fig:example-compilation}, S1 can receive advertisements in two different contexts $(S1,1,0)$ and $(S1,0,1)$ and must choose between messages received in these different contexts. Advertisements are preferred in state $(S1,1,0)$ because they result in a better path for S1 ($(S1,0,1) \geq_{rank} (S1,1,0)$) and downstream routers such as A1 will also obtain paths no worse than if S1 had choosen $(S1,1,0)$. The policy is guaranteed to be safe from failures because if a link fails in the topology, then $m \leq n$ will still hold since any transition that becomes unusable for $m$ also becomes unusable for $n$. That is $m \leq n$ before the failure implies $m \leq n$ after the failure.

For each router, its corresponding PG nodes are then sorted according the the $\leq$ operator. If the operator defines a total order on PG nodes, then the compiler can simply prefer advertisements from peers of node $m$ over those of $n$ whenever $m \leq n$. However, if $\leq$ does not form a total order, then the policy is rejected as being potentially unsafe under some failure conditions.

@@

Figure~\ref{fig:concretization} (right) defines compilation from \sysname to \mbgp. It proceeds by compiling constraints $p_i$ in the original \sysname policy to a tuple of: the predicate $t_i$, the product graph $PG_i$, and the preference function $pref_i$. These tuples are passed the the $\mathrm{compile}_\mathrm{mBGP}$ function, along with the network topology $G$. For each internal router in the topology $l$, and each predicate $t_i$ in the \sysname policy, compilation goes through each node $m$ for $l$ in $PG_i$, and groups the inbound neighbors of $m$ by tag ($q_n$) into sets ($in$). It allows imports from these neighbors before exporting to the outbound neighbors of $m$. The local preference for these imports is given by $\mathrm{pref}_i(m)$, which represents an integer based on the total ordering of $(\leq)$. We build configurations using list-comprehension notation. For instance, $[~ l \rightarrow rc ~\vert~ l \in V, p(l,rc) ~]$ denotes the \mbgp policy $l_1 \rightarrow rc_{1},...,l_k \rightarrow rc_{k}$ where each $rc_{i}$ satisfies $p(l_i,rc_{i})$. We use $\mathrm{append}$ to denote sequence concatenation.%

@@

\para{Concretization}
The similar structure between the spine template and concrete configurations is not a coincidence.  We formalize this observation by defining two concretization functions ($con$), one for \sysname policies and another for \mbgp policies. 
%
Concretization takes a context $\Gamma : \mathit{Var} \rightarrow 2^{\mathit{Prefix} \times V}$ that maps each template variable to a set of pairs of a concrete prefix and topology location where the prefix is owned.
%
Both concretization functions traverse the policy and substitute instances of a topology location $l$ in the template policy with the set of all concrete locations that map to $l$, given by the inverse homomorphism $f^{-1}(l) = \{l' ~\vert~ f(l') = l \}$. Additionally, whenever a pair $(pfx,l) \in \Gamma(x)$, then a new entry is added to the concretized policy where $pfx$ replaces $\$\mathrm{x}$ and adds the constraint that traffic ends at $l$ ($\CD{\End(l)}$).
%
For example, the spine template in Figure~\ref{fig:bgp-configs}, is obtained by substituting $\{ A1,A2 \}$ for $AG$ and $\{ A3,A4 \}$ for $AL$ and by also replacing the entry for {\small \$GP} with entries for {\small GP1} and {\small GP2} given by the context.
